---
title: "Mushroom classification in R"
output:
  html_document: default
  html_notebook: default
---

Using [Variable Selection](https://cran.r-project.org/web/packages/varSelRF/varSelRF.pdf) to label the mushroom into *p* (poisonous) and *e* (edible).

Load the data and quick check on it.
```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(varSelRF)
library(randomForest)

rm(list=ls())
my.dir <- "D:/Projects/5.0 Personal project/13.0 Mushroom/"
my.file <- "mushrooms.csv"
setwd(my.dir)

input <- paste0(my.dir, my.file)
raw.data <- read.csv(input, sep=",", header=T)
```

```{r}
head(raw.data, 1)
summary(raw.data)
str(raw.data)
```


Given we do not have separate testing and training dataset, we will divide the main datasets to training (70%) and testing (30%) dataset (with the help of createDataPartition function in caret library)

```{r, echo=F}
raw.data$class <- as.factor(raw.data$class)
print("=== Proportion of raw data set ===")
summary(raw.data$class)

# Data partition
training <- createDataPartition(y=raw.data$class, p=0.7, list=F)
training.data <- raw.data[training,]
testing.data <- raw.data[-training,]

print("=== Proportion of training data set ===")
summary(training.data$class)

print("=== Proportion of testing data set ===")
summary(testing.data$class)
```

Let's select the variables using varSelRF function
```{r}
# Using variable selection and variable importance
x <- dplyr::select(training.data, -class)
y <- as.factor(training.data$class)

var.sel <- varSelRF(x, y, ntree = 50, ntreeIterat = 20, vars.drop.frac = 0.3, whole.range = F, keep.forest = T)

# Initial importance variables
val.rank = var.sel$initialImportances[order(var.sel$initialImportances, decreasing =T)[1:ncol(x)], 1]
as.data.frame(val.rank)

varImpPlot(var.sel$rf.model)

plot(var.sel) 

plot(var.sel$rf.model, main="Error on # of tree")
```

Using only selected variables and the tree model for prediction
```{r, echo=T}

new.x <- subset(training.data, select=var.sel$selected.vars)
training.pred <- predict(var.sel$rf.model, new.x)
print("=== predict on training dataset ===")
confusionMatrix(training.pred, training.data$class, positive="p")


test.x <- subset(testing.data, select=var.sel$selected.vars)
test.pred <- predict(var.sel$rf.model, test.x)
print("=== predict on testing dataset ===")
confusionMatrix(test.pred, testing.data$class, positive="p")

```

This section is merely a playground ...
```{r, echo=F}
print(var.sel$selected.vars)
print(var.sel$rf.model)

model.imp <- importance(var.sel$rf.model)
plot(model.imp, ylab="MeanDecreaseGini")
impvar <- rownames(model.imp)[order(model.imp[, 1], decreasing = T)]

for(i in seq_along(impvar)){
  partialPlot(var.sel$rf.model, training.data[-1], impvar[i], xlab=impvar[i], 
              main=paste("Partial Dependence on", impvar[i]))
}
```






